---
title: "Lab 5A"
author: "Nehal Ur Rahman"
date: "`r Sys.Date()`"
output: word_document
---
NAME: NEHAL UR RAHMNAN
STUDENT ID: 991691259

Introduction: 
In lab 5A we are going to analyze Pokemon dataset and perform the given instructions to answer the questions.
This dataset is about the different types of pokemon and its strengths in terms of HP, Attack, Defense, Special attack, Special Defense and Speed.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#We are loading the below library to perform various functions.
```{r}
library(sets)
library(ggplot2)
library(tidyverse)
library(ISLR)
library(moments)
library(dplyr) 
library(GGally)
```
The Pokemon data is loaded by setting the session directory. 
```{r}
poke <- read.csv("Pokemon.csv")
head(poke)
```

We are using the cat function to print the column names of the pokemon dataset with the help of colnames function.
```{r}
cat("Column Names : ", colnames(poke))
```

Using Summary function`we display the statistical information about the numerical variables in the dataset.
```{r}
cat("\nSummary of data: \n \n")
summary(poke)
```

I have checked if there are any NA values present in the dataset to see if cleaning is required. 
```{r}
colSums(is.na(poke))
```

After going through the dataset I have noticed that there are some blank spaces in Type.2 variable. I have decided not to clean this variable as only some pokemon have its type 2 and thus it doesn't need to be cleaned. Cleaning it will lead to loss of a large chunk of data unnecessarily.

#Deleting any duplicates if present.
```{r}
poke <- poke[!duplicated(poke), ]
```


```{r}
str(poke)
```

The Pokemon dataset consists of 800 observations and 13 variables.
3 of the variables are of character type, 9 are of integer type and 1 of logical type.

##Question 1

Exploration
```{r}
#PLOT 1
barplot(table(poke$Type.1), main="Frequency of various Type 1 Pokemon", col="red" , xlab="Type 1 of Pokemon", ylab = "Frequency")
```

From the above barplot we can see the frequency of Type 1 pokemon in the dataset. The graph got is not a normal distribution graph and has many peaks.

```{r}
#PLOT 2
barplot(table(poke$Type.2), main="Frequency of various Type 2 Pokemon", col="skyblue" , xlab="Type 2 of Pokemon", ylab = "Frequency")
```

From the above barplot we can see the frequency of Type 2 pokemon in the dataset. The graph got is skewed towards the right and is uni-modal. The plot is unimodal as it has only one peak



```{r}
#PLOT 3
as.factor(poke$Legendary)
barplot(table(poke$Legendary), main="Total number of Legendary Pokemon", col="gold" , xlab="TRUE = LEGENDARY POKEMON & FALSE = NOT LEGENDARY POKEMON")
```

The plot shown above gives us information of how many legendary pokemons are there in the dataset. Therefore, most of the pokemons are not legendary in the given dataset and a very small number of pokemons are legendary.

```{r}
#PLOT 4
ggcorr(poke, nbreaks = 6, label = TRUE, label_size = 3)
```

With the help correlation matrix given above, we can see how different numerical variables are strongly or weakly related to each other.
From the correlation matrix we get to know that the TOTAL variable has strong correlation with HP, Attack, Defense, Special attack, Special Defense and Speed.
The variable X and Generation are strongly correlated as well.

##Question 2

#Step 1
First we create another variable called HP_category which replicates the values of HP.
The HP_category is the stored in the poke dataset along with other variables. So now we have 14 variables.
```{r}
HP_category <- rep(poke$HP ,)
poke <- data.frame(poke , HP_category)
```

```{r}
str(poke)
```


#Step 2
Now, we have categorized the Pokemon based on its HP. Therefore, 6 levels of HP_category have been created for the given sets of HP using ifelse condition.
```{r}
poke$HP_category<-  as.factor(ifelse(poke$HP_category <= 49, 'Level 1: (-∞, 49]', 
                    ifelse(poke$HP_category >= 50 & poke$HP_category <= 99,'Level 2: [50, 99]',
                    ifelse(poke$HP_category >= 100 & poke$HP_category <= 149,'Level 3: [100, 149]',
                    ifelse(poke$HP_category >= 150 & poke$HP_category <= 199, 'Level 4: [150, 199]',
                    ifelse(poke$HP_category >= 200 & poke$HP_category <= 249, 'Level 5: [200, 249]',
'Level 6: [250, +∞)'))))))

```

##Question 3

Using tapply function the average Attack strength for each HP_category is found.
```{r}
tapply(poke$Attack, poke$HP_category, FUN = mean)
```
From the above output we camn infer that the mean attack strength is the highest for Level 3 Pokemon and the least for level 6 pokemon


##Question 4

#Step 1 
First we see the total number of legendary pokemon in the dataset(TRUE = LEGENDARY POKEMON & FALSE = NOT LEGENDARY POKEMON)
```{r}
table(poke$Legendary)
```
There are total of 65 legendary pokemon and 735 non-legendary pokemon in the dataset.

#Step 2
Now we check if Mega is present in the name of the Pokemon using grepl function and store its value in the poke dataset.
```{r}
mega_poke<-grepl("Mega",poke$Name)
poke<-data.frame(poke,mega_poke)
table(mega_poke)
```
The grepl function returns a logical value(TRUE or FALSE) depending on the word pattern given.
Therefore, there are a total of 49 Mega-Evolved Pokémon and 751 non-Mega-Evolved Pokémon.

#Step 3 
Now we make another dataset as the subset of poke in which there are only legendary pokemons.
```{r}
legendary_poke <- subset(poke,poke$Legendary=='TRUE')
str(legendary_poke)
```

Now we have 15 variables in the dataset after adding the variable mega_poke.

#Step 4
We substitute the True with Mega Poke and False with Non Mega to get a better idea of the proportion.
```{r}
legendary_poke$mega_poke<-gsub("TRUE","Mega Poke",legendary_poke$mega_poke)
legendary_poke$mega_poke<-gsub("FALSE","NON-Mega Poke",legendary_poke$mega_poke)
```

Using the proportion table we are able to find out proportion of the Mega-Evolved Pokémon and non-Mega-Evolved Pokémon that have a Legendary value(Legendary = TRUE). 
```{r}
prop.table(table(legendary_poke$mega_poke))
```

With the output above, we  understand that the proportion of Mega-Evolved Pokémon and non-Mega-Evolved Pokémon are: 
Mega-Evolved Pokémon = 9.23%
Non-Mega-Evolved Pokémon = 90.76% 
