---
title: "Lab 5A"
author: "Nehal Ur Rahman"
date: "`r Sys.Date()`"
output: word_document
---
NAME: NEHAL UR RAHMNAN
STUDENT ID: 991691259

Introduction: 
In lab 5A we are going to analyze Pokemon dataset and perform the given instructions to answer the questions.
This dataset is about the different types of pokemon and its strengths in terms of HP, Attack, Defense, Special attack, Special Defense and Speed.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#We are loading the below library to perform various functions.
```{r}
library(sets)
library(ggplot2)
library(tidyverse)
library(ISLR)
library(moments)
library(dplyr) 
library(GGally)
```
The Pokemon data is loaded by setting the session directory. 
```{r}
poke <- read.csv("Pokemon.csv")
head(poke)
```

We are using the cat function to print the column names of the pokemon dataset with the help of colnames function.
```{r}
cat("Column Names : ", colnames(poke))
```

Using Summary function`we display the statistical information about the numerical variables in the dataset.
```{r}
cat("\nSummary of data: \n \n")
summary(poke)
```

I have checked if there are any NA values present in the dataset to see if cleaning is required.
```{r}
colSums(is.na(poke))
```
```{r}
str(poke)
```

The Pokemon dataset consists of 800 observations and 13 variables.
3 of the variables are of character type, 9 are of integer type and 1 of logical type.

##QUestion 1

Exploration
```{r}
barplot(table(poke$Type.1), main="Barplot of Different Type 1 Pokeomon", col="red" , xlab="Type 1 of Pokemon ")
```

```{r}
barplot(table(poke$Type.2), main="Barplot of Different Type 2 Pokeomon", col="skyblue" , xlab="Type 2 of Pokemon")
```

```{r}
as.factor(poke$Legendary)
barplot(table(poke$Legendary), main="Barplot of total number of Legendary Pokemon", col="gold" , xlab="TRUE = LEGENDARY POKE & FALSE = NOT LEGENDARY POKE")
```
```{r}
ggcorr(poke, nbreaks = 6, label = TRUE, label_size = 3)
```

```{r}
HP_category <- rep(poke$HP ,nrow(poke))
poke <- data.frame(poke , HP_category)
```


```{r}
poke$HP_category<-  as.factor(ifelse(poke$HP_category <= 49, 'Level 1', 
                    ifelse(poke$HP_category >= 50 & poke$HP_category <= 99,'Level 2',
                    ifelse(poke$HP_category >= 100 & poke$HP_category <= 149,'Level 3',
                    ifelse(poke$HP_category >= 150 & poke$HP_category <= 199, 'Level 4',
                    ifelse(poke$HP_category >= 200 & poke$HP_category <= 249, 'Level 5',
'Level 6'))))))

```


```{r}
tapply(poke$Attack, poke$HP_category, FUN = mean)
```
```{r}
table(poke$Legendary)
```
```{r}
mega_poke<-grepl("Mega",poke$Name)
poke<-data.frame(poke,mega_poke)
table(mega_poke)
```
```{r}
legendary_poke <- subset(poke,poke$Legendary=='TRUE')
```

```{r}
legendary_poke$mega_poke<-gsub("TRUE","Mega Poke",legendary_poke$mega_poke)
legendary_poke$mega_poke<-gsub("FALSE","NON-Mega Poke",legendary_poke$mega_poke)
```


```{r}
prop.table(table(legendary_poke$mega_poke))
```

